---
title: "Info8000 Assignment 3"
output: html_document
---
Description: Exploratory visualization of survey results

Questions:

Q1. What is the distribution of demographic characteristics of the respondents?
    Income, Ethnicity, and Education variables visualized as bar graphs:
```{r}
library(tidyverse)
sgeo <- read.csv("~/Desktop/sgeoTable.csv")
a <- ggplot(data = sgeo)

a +
  geom_bar(mapping = aes(x = as.factor(Hincome))) +
    scale_x_discrete(breaks = c(1:7), 
                     labels = c("<25K","25-35K","35-50K","50-75K","75-100K","100-150K",">150K")) +
  labs(x = "Income Category")
a +
  geom_bar(mapping = aes(x = as.factor(Ethnicity))) +
    scale_x_discrete(breaks = c(1:8), 
                  labels = c("White","Hispanic","Black","Asian","Native American",
                                "MidEast/N.Africa","Pacific Islander","Other")) +
  labs(x = "Ethnicity")
a +
  geom_bar(mapping = aes(x = as.factor(Education))) +
  scale_x_discrete(breaks = c(1:7),
                   labels = c("Did not finish HS","HS/GED","Some college","Assoc","Bachelors","Some grad","Grad degree")) +
  labs(x = "Educational Attainment")
```

Q2. Do respondents in high-risk flood zones express greater concern about severe storms?

Q3. Do income, race, or ethnicity relate to how respondents perceive their ability to recover from storms?
    Less educated and lower income individuals, and members of minority communities are theorized to be 
    less able to recover from the impacts of severe storms. These graphs break down respondents perceived ability
    to recover from storm damage based on these variables. The results are represented as stacked bar graphs and as
    boxplots.

```{r}
  #by ethnicity
ggplot(sgeo, aes(PCNoRecov))+
  geom_bar(aes(fill = as.factor(Ethnicity))) +
  scale_fill_discrete(name = "Ethnicity", breaks = c(1:8),
                      labels = c("White","Hispanic","Black","Asian","Native American",
                                 "MidEast/N.Africa","Pacific Islander","Other")) +
  labs(x = "Ability to recover")
a +
  geom_boxplot(mapping = aes(x = as.factor(Ethnicity),y = PCNoRecov))+
  scale_x_discrete(breaks = c(1:8), 
                   labels = c("White","Hispanic","Black","Asian","Native American",
                              "MidEast/N.Africa","Pacific Islander","Other"))+
  labs(x = "Ethnicity",y = "Ability to recover")
  #by income
ggplot(sgeo, aes(PCNoRecov))+
  geom_bar(aes(fill = as.factor(Hincome))) +
  scale_fill_discrete(name = "Income Category", breaks = c(1:7),
                      labels = c("<25K","25-35K","35-50K","50-75K","75-100K","100-150K",">150K")) +
  labs(x = "Ability to recover")
a+ 
  geom_boxplot(mapping = aes(x = as.factor(Hincome), y = PCNoRecov))+
  scale_x_discrete(breaks = c(1:7), 
                   labels = c("<25K","25-35K","35-50K","50-75K","75-100K","100-150K",">150K")) +
  labs(x = "Income Category", y = "Ability to recover")

  #by education
ggplot(sgeo, aes(PCNoRecov))+
  geom_bar(aes(fill = as.factor(Education))) +
  scale_fill_discrete(name = "Educational attainment", breaks = c(1:7),
                      labels = c("Did not finish HS","HS/GED","Some college","Assoc","Bachelors","Some grad","Grad degree")) +
  labs(x = "Ability to recover")
a +
  geom_boxplot(mapping = aes(x = as.factor(Education),y = PCNoRecov)) +
  scale_x_discrete(breaks = c(1:7),
                   labels = c("Did not finish HS","HS/GED","Some college","Assoc","Bachelors","Some grad","Grad degree"))+
  labs(x = "Educational Attainment", y = "Ability to recover")
```


